Climate indicators

### 2023-03-07 DRAFT_climate_indicators.Rmd

The reference condition for climate variables is defined as the last normal period 1961-1990.

@@

-   Un-scaled indicator value = median value over 5 year periods

@@

>>>>

The reference condition for climate variables is defined as the normal period 1961-1990.

@@

-   Un-scaled indicator value = median value over 5 year periods (5 years being a pragmatic choice. It is long enough to smooth out a lot of inter-annual variation, and it's long enough to enable estimating errors)

@@

The choice to use 2 SD units as the threshold values is a subjective choice in many ways, and not founded in any empirical or known relationship between the indicators and ecosystem integrity. The value comes from the common practice of calling something _extreme weather_ when it is outside 2 SD units of the current running average. So, if the indicator value today is <0.6 it implies that the mean for that variable over the last year would have been referred to as _extreme_ if it had occurred between 1961 and 1990. This is I think a conservative threshold, since one would/could call it _extreme_ if only one year is outside the 2SD range, and having the mean of 5 years being outside this range is _really_ extreme.
